Spin canonical divisors of log stacky curves

### 2015-06-21 stacky_write_up/spin_cring.tex
reintroduce e_i - e_j', index base cases
added more proof to section 3 main lemma

\newcommand \subhalf[1]{\frac{{#1} - 1}{2{#1}}}
\newcommand{\se}[1]{\section*{Problem #1}}
\newcommand \halfcan{L}
\newcommand \Supp{\text{Supp}}
\newcommand \initial{\text{in}}

@@

\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{dD + d\frac{\alpha}{\beta} P \rfloor}) = h^0(X,\lfloor dD
	\rfloor) + \lfloor d\frac{\alpha} {\beta} \rfloor \forall d \in \mathbf{
	N} : d \ge \frac{\beta}{\alpha}.
\end{equation}
Let 
\[
	0<\frac{c_1}{d_1}<...<\frac{c_n}{d_n}=\frac{\alpha}{\beta}
\]
be the non-negative best lower approximation.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1,...,y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$
\item If $I$ and $I'$ are the ideal of relations of $R_D$ and $R_{D'}$ respectively, then

\begin{align*}
	\initial_\prec(I') &= \initial_\prec(I) k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \\
										 &+ \langle y_i x_j: 1 \le i \le n, 1 \le j \le m \rangle \\
										 &+ \langle y_i y_j: 1 \le i \le j \le n-1 \rangle.
\end{align*}

\todo{See if we can additionally get the other information discussed in VZB theorem 8.3.1}
\end{itemize}
\end{lem}

@@

Let $0 < \frac{c_1}{d_1} < \ldots < \frac{c_n}{d_n} = \frac{\alpha}{
\beta}$ be the non-negative best lower approximation of $\frac{
\alpha}{\beta}$.  Note that elements of $H^0(X,dD')$ have pole of
orders ranging from 0 to $\lfloor d_i \frac{\alpha}{\beta}\rfloor =
c_i $ at $P$. A set of these one for each pole degree 1 to $\lfloor
d_i \frac{\alpha}{\beta} \rfloor$ are linearly independent from
each other and $(R_D)_d$, so by degree considerations, they span
$(R_{D'})_d$ over $(R_D)_d$ as $k$-vector spaces.

For any $j < i$, $\frac{c_i}{d_i} > \frac{c_j}{d_j}$
by construction, so any product of $\prod z_i$ of elements of
degrees less than $d_i$ must have $\frac{-\ord(z_i)}{\deg(z_i)} <
\frac{c_i}{d _i}$, so we have the pole order of this product is $
\sum -\ord(z_i) = d_i(\frac{\sum -\ord( z_i)}{\sum \deg( z_i)}) < d_
i\frac{c_i}{d_i} = c_i $. Thus, the elements of $R_{d_i}$

For any $j<i$, $\frac{c_i}{d_i}>\frac{c_j}{d_j}$ by construction,
so any product of $\prod z_i$ of elements of degrees less than $d_i$
 must have $\frac{-\ord(z_i)}{\deg(z_i)}<\frac{c_i}{d_i} $, so we
have the pole order of this product is $\sum -\ord(z_i)=d_i(\frac{
\sum -\ord( z_i)} {\sum \deg( z_i)})<d_i\frac{c_i}{d_i}=c_i$.
Thus, elements of $(R_{D'})d_i$ with pole order $c_i$ at $P$ are
not generated by lower degrees; generically choose some $y_i \in(R_{
D'})_{d_i}$ to have maximal pole order (note that if $d_i=1$ then $-
\ord_P(y_i)= \max(c_j:d_j=1)$ and if more than one $d_i$ is 1, we
must also choose them generically so they generate $(R_{D'})_1$).

Suppose $d\le \beta$ and $\frac{\bar{c}}{\bar{d}} \le \frac{\alpha}{
\beta}$ with $c>0$ is not a best lower approximation.  Then choose
best lower approximation $\frac{c_i}{d_i}$ such that $c_i\le c$ is
maximal; since $c<c_{i+1}$, $\frac{c_i}{d_i}\ge \frac{c}{d}$.  From
this we can deduce that $c_i+(d-d_i)\lfloor \frac{\alpha}{\beta}
\rfloor\ge c$.  Therefore, we can produce an element in $(R_{D'})_d$
 with pole degree $c$ at $P$ by either $y_i u^b$ or $z y_i u^b$ with $i>\lfloor \frac{\alpha}{\beta} \rfloor$, or $zu^b$
for some $z=y_j (y_{\lfloor \frac{\alpha}{\beta} \rfloor})^b$ where $d_j=1$ ($=d_{{\lfloor \frac{a}{b} \rfloor}}$) and $b<d_{\lfloor \frac{\alpha}{\beta} \rfloor + 1}$.  Thus, for a
general $d$ (not necessarily less than $\beta$), if $\frac{c}{d} \le
\frac{\alpha}{\beta}$ is not a best lower approximation than we can
construct an element in degree $d$ with pole degree $c$ by either $y
_n^a y_i u^b$, $y_n^a zy_iu^b$ or $y_n^a z u^b$ satisfying the
same conditions.  We notice that elements of this form are all
linearly independent in $R_{D'}$.  Thus, they form a $k$-basis for
$R_{D'}$ over $R_D$.  Thus $u, x_1, ..., x_m, y_1, ..., y_n$ generate
$R_{D'}$.

Choose $x_i\in \{x_1, \ldots, x_m\}$ and $y_j\in \{y_1, \ldots, y_n\}$, where $y_j$ corresponds to the best lower approximation $\frac{c_j}{d_j}$.  Recall from the previous paragraph that we can uniquely construct an element in degree $d$ with pole order $c$ at $P$ by elements of the $y_n^a y_l u^b$ and $y_n^a y_l y_{l'} u^b$ with $d_l=1$ and if $d_{l'}=1$ then $c_{l'} = \lfloor \frac{\alpha}{\beta} \rfloor$.  Specifically, we can successively subtract some $\gamma_l y_l u^{b_l}$, $\gamma_l y_l z_l u^{b_l}$, or $\gamma_l z_l$ with $y_l, z_l, b_l$ satisfying the same conditions as above and $\gamma_l\in k$ the element ensuring that the difference decreases in pole degree at $P$ by 1.  For convenience, we can combine these into the case $\gamma_l (y_l)^{s_l}(z_l)^{a_l}u^{b_l}$ with $(s_l,a_l)\in \{(1,0),(1,1),(0,1)\}$.  In the cases when $\frac{\alpha}{\beta}<1$ so $z_l$ is not well-defined, just pick an arbitrary element of $R_D$ for $z_l$ ensuring that the expression still makes sense; in such cases we will always choose $a_l=0$ so this choice is irrelevant, and only makes writing out cases more convenient.  Therefore, we can write
\[
	x_i y_j - \sum_{l} (y_l)^{s_l} (z_l)^{a_l} u^{b_l}\in R_D
\]
giving us a non-trivial relation with leading term $x_iy_j$. \todo{Explicitly state ordering at the beginning}

Similarly, suppose $y_i,y_j\in \{y_1, \ldots ,y_{n-1}\}$ with $1<d_i$ and $j\ge i$ (that is, so $y_iy_j$ does not appear as a basis element of the form described above: $y_n^0 y_jy_i u^0$).  Choose the maximal $l$ such that $c_h\le c_i+c_j$; since $1<d_i\le d_j$, $(c_i+c_j)-c_j>\lfloor \frac{\alpha}{\beta}\rfloor$ so $c_{j+1} < c_i + c_j$ meaning $c_h \ge c_{j+1} > c_j$.  Furthermore, $c_i + c_j - c_h<\lfloor \frac{\alpha}{\beta} \rfloor$ so either $c_h=c_i+c_j$ or there is some $h'$ such that $c_{h'}=c_i+c_j-c_l$ and $d_{h'}=1$.  Since $\frac{c_h}{d_h}$ is a best lower approximation of $\frac{\alpha}{\beta}$ with denominator larger than those of $\frac{c_i}{d_i}$ and $\frac{c_j}{d_j}$, we know that since $c_h\le c_i+c_j$ we must have $d_h<d_i+d_j$.  Thus in the case with a $y_{l'}$ term we have $d_h + d_{h'} = d_h + 1 \le d_i + d_j$.  That is to say, by choosing $b-d_i+d_j-ad_h-1$ we find (where $a\in\{0,1\}$ corresponding to if a $y_{h'}$ appears) that $y_h(y_h')^au^b$ and $y_iy_j$ both lie in degree $d_i+d_j$ with pole order $c_i+c_j$ at $P$.  From here we can follow a similar technique as in the previous paragraph to cancel out poles of $y_iy_j$ at $P$ to find
\[
	y_iy_j-\gamma_h (y_h)^{s_h}(z_{h})^{a_h}u^{b_h}-\sum_l (y_l)^{s_l} (z_{l})^{a_l}u^{b_l}\in R_D
\]
implying a relation with initial term $y_i y_j$.

Finally, suppose $y_i, y_j\in \{y_1, \ldots ,y_{n-1}\}$ such that $i\le j$, $d_j=d_i=1$, and $d_{j+1}=1$ (i.e. $c_i, c_j \le \lfloor \frac{\alpha}{\beta} - 1 \rfloor$).  If $i>1$ and then $y_{j+1}y_{i-1}$ lies in the same degree as $y_iy_j$ (degree 2) with the same pole order; following the same method as the previous two paragraphs we find
\[
	y_i y_j - \gamma y_{i-1} y_{j+1} - \sum_{l} (y_l)^{s_l} (z_l)^{a_l}u^{b_l}\in R_D.
\]
Similarly if $i=1$, then we get the same result using
\[
	y_i y_j - \gamma u y_{j+1} - \sum_l y_l u^{b_l} \in R_D.
\]
In both cases we deduce that $y_iy_j$ is an initial term of some relation.

Finally, we examine the case (for $\alpha \ge \beta$) of $(y_i)^{i + 1}$ when $i=\lfloor \frac{\alpha}{\beta}\rfloor$.  If $\beta=1$, then there are no relations with initial term a power of $y_i$ and we are immediately done.  Otherwise, assuming $\beta\ne 1$, we note that $c_id_{i+1}=c_{i+1}d_i+1=c_{i+1}+1$. \todo{reference to Evan's paper where he proves this}  Therefore, if $i>1$ then $c_i^{d_{i+1}+1}$ and $c_{i-1} c_{i+1}$ both lie in the $d_{i+1}+1$ degree with pole order $c_{i+1}+1$.  Following a similar technique to the previous paragraphs we find
\[
	(y_{i})^{c+1} - y_{i+1} y_{i-1} - \sum_l (y_l)^{s_l}(z_l)^{a_l} u^{b_l}
\]

Let $J$ be the ideal of $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ generated by the initial terms described in the previous paragraphs.  We show that $J$ is in fact all of the initial ideal.  Suppose $f$ is a non-constant monomial that does not lie within $J$.  Further suppose $f$ does not lie in $k[u, x_1, \ldots, x_m]$.  Then there is some $y_i|f$.  Since $f\not\in j$, we cannot have some $x_j|f$ or else we would have $y_ix_j|f$ contradicting $f\not\in J$.  Furthermore, if $y_l|f$ for $l<i$, then we must have $d_l=1$.  If $d_i=1$ as well, then we must have $c_i=\lfloor \frac{\alpha}{\beta}\rfloor$, or else.  We observe, that such elements are precisely

\end{proof}

\todo{lots of corollaries; peter can do this tomorrow}

>>>>

\newcommand \subhalf[1]{\frac{{#1} - 1}{2{#1}}}
\newcommand{\se}[1]{\section*{Problem #1}}
\newcommand \halfcan{L}
\newcommand \Supp{\text{Supp}}
\newcommand \initial{\text{in}}
\newcommand \gin{\text{gin}}

@@

\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{dD + d\frac{\alpha}{\beta} P \rfloor}) = h^0(X,\lfloor dD
	\rfloor) + \lfloor d\frac{\alpha} {\beta} \rfloor \forall d \in \mathbf{
	N} : d \ge \frac{\beta}{\alpha}.
\end{equation}
Let 
\[
	0<\frac{c_1}{d_1}<...<\frac{c_n}{d_n}=\frac{\alpha}{\beta}
\]
be the non-negative best lower approximation.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1,...,y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$

\item If $I$ and $I'$ are the ideal of relations of $R_D$ and $R_{D'}$ respectively, then

\begin{align*}
	\initial_\prec(I') &= \initial_\prec(I) k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \\
										 &+ \langle y_i x_j: 1 \le i \le n, 1 \le j \le m \rangle \\
										 &+ \langle y_i y_j: \lfloor \frac{\alpha}{\beta} \rfloor \rangle \le i \le j \le n-1 \\
										 &+ \langle y_i y_j: 1\le i\le j\lfloor \frac{\alpha}{\beta} \rfloor\\
										 &+ \langle (y_i)^{\deg(y_{i+1})}: 1\le i=\lfloor \frac{\alpha}{\beta} \rfloor
\end{align*}
In particular, if $\frac{\alpha}{\beta}\le 1$, then $I$ is minimally generated by elements with leading terms of the form $y_ix_j$ and $y_iy_j$.
\item If $\gin_\prec(I)=\initial_\prec(I)$ then $\gin_\prec(I')=\initial_\prec(I')$.
\end{itemize}
\end{lem}

@@

Let 
\[
	0 < \frac{c_1}{d_1} < \ldots < \frac{c_n}{d_n} = \frac{\alpha}{
\beta}
\] 
be the non-negative best lower approximation of $\frac{
\alpha}{\beta}$.  Note that elements of $H^0(X,dD')$ have pole of
orders ranging from 0 to $\lfloor d_i \frac{\alpha}{\beta}\rfloor =
c_i $ at $P$.  A set of these one for each pole degree 1 to $\lfloor
d_i \frac{\alpha}{\beta} \rfloor$ are linearly independent from
each other and $(R_D)_d$, so by degree considerations, they span
$(R_{D'})_d$ over $(R_D)_d$ as $k$-vector spaces.

For any $j < i$, $\frac{c_i}{d_i} > \frac{c_j}{d_j}$
by construction, so any product of $\prod z_i$ of elements of
degrees less than $d_i$ must have 
\[
	\frac{-\ord(z_i)}{\deg(z_i)} < \frac{c_i}{d _i},
\]
so we have the pole order of $\prod z_i$ is 
\[
	\sum -\ord(z_i) = d_i(\frac{\sum -\ord( z_i)}{\sum \deg( z_i)}) < d_i\frac{c_i}{d_i} = c_i .
\] 
Thus, the elements of $(R_{D'})_{d_i}$ with pole order $c_i$ at $P$ are not generated by 
lower degrees.  
Generically choose some $y_i \in(R_{
D'})_{d_i}$ to have maximal pole order at $P$ (note that if $d_i=1$ then 
\[
	-\ord_P(y_i)= \max(c_j:d_j=1)
\]
and if more than one $d_i$ is 1, we
must also choose them generically so they generate $(R_{D'})_1$).

Suppose $d\le \beta$ and positive $\frac{c}{d} \le \frac{\alpha}{
\beta}$ not a best lower approximation.  Then choose
best lower approximation $\frac{c_i}{d_i}$ such that $c_i\le c$ is
maximal; since $c<c_{i+1}$, $\frac{c_i}{d_i}\ge \frac{c}{d}$.  From
this we can deduce that 
\[
	c_i+(d-d_i) \lfloor \frac{\alpha}{\beta} \rfloor \ge c.
\]  
Therefore, we can generate an element of $(R_{D'})_d$ as a product of lower degrees that 
has pole degree $c$ at $P$ by either $y_i u^b$ or $y_i z u^b$, or $zu^b$ satisfying $i>
\lfloor \frac{\alpha}{\beta} \rfloor$ and $z=y_j (y_{\lfloor \frac{\alpha}{\beta} \rfloor})^b$ 
where $d_j=1$ ($=d_{{\lfloor \frac{a}{b} \rfloor}}$) and $b<d_{\lfloor \frac{\alpha}{\beta} 
\rfloor + 1}$.  
Thus, for a general $d$ (not necessarily less than $\beta$), if $\frac{c}{d} \le
\frac{\alpha}{\beta}$ is not a best lower approximation than we can
construct an element in degree $d$ with pole degree $c$ by either $y
_n^a y_i u^b$, $y_n^a zy_iu^b$ or $y_n^a z u^b$ satisfying the
same conditions.  We notice that elements of this form are all
linearly independent in $R_{D'}$.  Thus, they form a $k$-basis for
$R_{D'}$ over $R_D$.  Thus $u, x_1, ..., x_m, y_1, ..., y_n$ generate
$R_{D'}$.

Choose $x_i\in \{x_1, \ldots, x_m\}$ and $y_j\in \{y_1, \ldots, y_n\}$, where $y_j$ 
corresponds to the best lower approximation $\frac{c_j}{d_j}$.  Recall from the previous 
paragraph that we can uniquely construct an element in degree $d$ with pole order $c$ at 
$P$ by elements of the $y_n^a y_l u^b$ and $y_n^a y_l y_{l'} u^b$ with $d_l=1$ and if 
$d_{l'}=1$ then $c_{l'} = \lfloor \frac{\alpha}{\beta} \rfloor$.  Specifically, we can 
successively subtract some $\gamma_l y_l u^{b_l}$, $\gamma_l y_l z_l u^{b_l}$, or $
\gamma_l z_l$ with $y_l, z_l, b_l$ satisfying the same conditions as above and $\gamma_l
\in k$ the element ensuring that the difference decreases in pole degree at $P$ by 1.  For 
convenience, we can combine these into the case $\gamma_l (y_l)^{s_l}(z_l)^{a_l}u^{b_l}$ 
with $(s_l,a_l)\in \{(1,0),(1,1),(0,1)\}$.  In the cases when $\frac{\alpha}{\beta}<1$ so $z_l$ 
is not well-defined, just pick an arbitrary element of $R_D$ for $z_l$ ensuring that the 
expression still makes sense; in such cases we will always choose $a_l=0$ so this choice 
is irrelevant, and only makes writing out cases more convenient.  Therefore, we can write
\[
	x_i y_j - \sum_{l} (y_l)^{s_l} (z_l)^{a_l} u^{b_l}\in R_D
\]
giving us a non-trivial relation with leading term $x_iy_j$. \todo{Explicitly state ordering at 
the beginning}

Similarly, suppose $y_i,y_j\in \{y_1, \ldots ,y_{n-1}\}$ with $1<d_i$ and $j\ge i$ (that is, so 
$y_iy_j$ does not appear as a basis element of the form described above: $y_n^0 y_jy_i 
u^0$).  Choose the maximal $l$ such that $c_h\le c_i+c_j$; since $1<d_i\le d_j$, $(c_i
+c_j)-c_j>\lfloor \frac{\alpha}{\beta}\rfloor$ so $c_{j+1} < c_i + c_j$ meaning $c_h \ge c_{j
+1} > c_j$.  Furthermore, $c_i + c_j - c_h<\lfloor \frac{\alpha}{\beta} \rfloor$ so either 
$c_h=c_i+c_j$ or there is some $h'$ such that $c_{h'}=c_i+c_j-c_l$ and $d_{h'}=1$.  Since 
$\frac{c_h}{d_h}$ is a best lower approximation of $\frac{\alpha}{\beta}$ with denominator 
larger than those of $\frac{c_i}{d_i}$ and $\frac{c_j}{d_j}$, we know that since $c_h\le c_i
+c_j$ we must have $d_h<d_i+d_j$.  Thus in the case with a $y_{l'}$ term we have $d_h + 
d_{h'} = d_h + 1 \le d_i + d_j$.  That is to say, by choosing $b-d_i+d_j-ad_h-1$ we find 
(where $a\in\{0,1\}$ corresponding to if a $y_{h'}$ appears) that $y_h(y_h')^au^b$ and 
$y_iy_j$ both lie in degree $d_i+d_j$ with pole order $c_i+c_j$ at $P$.  From here we can 
follow a similar technique as in the previous paragraph to cancel out poles of $y_iy_j$ at $P
$ to find
\[
	y_iy_j-\gamma_h (y_h)^{s_h}(z_{h})^{a_h}u^{b_h}-\sum_l (y_l)^{s_l} (z_{l})^{a_l}
u^{b_l}\in R_D
\]
implying a relation with initial term $y_i y_j$.

Next, suppose $y_i, y_j\in \{y_1, \ldots ,y_{n-1}\}$ such that $i\le j$, $d_j=d_i=1$, and $d_{j
+1}=1$ (i.e. $c_i, c_j \le \lfloor \frac{\alpha}{\beta} - 1 \rfloor$).  If $i>1$ and then $y_{j+1}
y_{i-1}$ lies in the same degree as $y_iy_j$ (degree 2) with the same pole order; following 
the same method as the previous two paragraphs we find
\[
	y_i y_j - \gamma y_{i-1} y_{j+1} - \sum_{l} (y_l)^{s_l} (z_l)^{a_l}u^{b_l}\in R_D.
\]
Similarly if $i=1$, then we get the same result using
\[
	y_i y_j - \gamma u y_{j+1} - \sum_l y_l u^{b_l} \in R_D.
\]
In both cases we deduce that $y_iy_j$ is an initial term of some relation.

Finally, we examine the case (for $\alpha \ge \beta$) of $(y_i)^{i + 1}$ when $i=\lfloor 
\frac{\alpha}{\beta}\rfloor$.  If $\beta=1$, then there are no relations with initial term a 
power of $y_i$ and we are immediately done.  Otherwise, assuming $\beta\ne 1$, we note 
that $c_id_{i+1}=c_{i+1}d_i+1=c_{i+1}+1$. \todo{reference to Evan's paper where he 
proves this}  Therefore, if $i>1$ then $c_i^{d_{i+1}+1}$ and $c_{i-1} c_{i+1}$ both lie in the 
$d_{i+1}+1$ degree with pole order $c_{i+1}+1$.  Following a similar technique to the 
previous paragraphs we find
\[
	(y_i)^{c_{i+1}+1} - y_{i+1} y_{i-1} - \sum_l (y_l)^{s_l}(z_l)^{a_l} u^{b_l}.
\]
Similarly if $i=1$ then
\[
	(y_i)^{c_{i+1}+1}-y_{i+1}u-\sum_l (z_l)^{a_l} u^{b_l}.
\]
In each case we demonstrate $(y_i)^{c_{i+1}+1}$ as the initial term of a relation.

Let $J'$ be the ideal of $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ generated by the initial 
terms described in the previous paragraphs together with $\gin_\prec(I)k[u, x_1, \ldots, 
x_m, y_1, \ldots, y_n]$.  We show that $J$ is in fact all of the initial ideal.  Suppose $f$ is a 
non-constant monomial that does not lie within $J$.  Further suppose $f$ does not lie in 
$k[u, x_1, \ldots, x_m]$.  Then there is some $y_i|f$.  Since $f\not\in j$, we cannot have 
some $x_j|f$ or else we would have $y_ix_j|f$ contradicting $f\not\in J$.  Furthermore, if 
$y_l|f$ for $l<i$, then we must have $d_l=1$ or else $y_ly_i$ is an initial element of a 
relation and divides $f$.  If $d_i=1$ as well, then we must have $c_i=\lfloor \frac{\alpha}
{\beta}\rfloor$, or else $y_iy_l\in I$; finally we can not have $(y_{\lfloor \frac{\alpha}{\beta}
\rfloor})^{d_{\lfloor \frac{\alpha}{\beta} \rfloor+1}+1}|f$.  We observe that the remaining 
elements are precisely those of the form $y_n^a y_i u^b$, $y_n^a zy_iu^b$, and $y_n^a z 
u^b$ which are the elements of our chosen $k$-basis for $k[u, x_1, \ldots, x_m, y_1, \ldots, 
y_n]/I'$ over $k[u, x_1, \ldots, x_m]/I$.  Therefore, by dimension counting (in each degree), 
we see that $J'$ is the initial ideal of $I'$.  Furthermore, since our choice of $y_i$s was 
generic, this is in fact the generic initial ideal.
\end{proof}

This yields some immediate corollaries:

\begin{cor}

\end{cor}
